%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Problem Set/Assignment Template to be used by the
%% Food and Resource Economics Department - IFAS
%% University of Florida's graduates.
%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Version 1.0 - November 2019
%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Ariel Soto-Caro
%%  - asotocaro@ufl.edu
%%  - arielsotocaro@gmail.com
%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Numbered environment




\documentclass[12pt]{article}
\usepackage{design_ASC}
\theoremstyle{definition}
\newtheorem{exmp}{Example}[section]
\newtheorem{slo}{Solution}[section]
\newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
\newcommand*{\Comb}[2]{{}^{#1}C_{#2}}%
%% -----------------------------
%% TITLE
%% -----------------------------
\title{\textbf{Lecture 5}} %% Assignment Title

\author{\textbf{Ibrahim} Abou Elenein}

\date{\today} %% Change "\today" by another date manually
%% -----------------------------
%% -----------------------------
% %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\setlength{\droptitle}{-5em}    
%% %%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle
\tableofcontents
\pagebreak
% --------------------------
% Start here
% --------------------------
\section{Conditional Probability}
The conditional probability of an event
A occurring given an event B “i.e.,
knowing that B has already occurred
first” is given by:
$$P(A|B) = \frac{P(A \cap B)}{P(B)}$$

\section{Independent Events}
In general, two events A and B are called independent if 
$$ P(A \cap B) = P(A) \times P(B)$$
So we can say that 
$$ P(A|B) = P(A) $$
$$ P(B|A) = P(B) $$
Also $$ P(A\cap B \cap C) = P(A)\times P(B) \times P(C)$$

\begin{exmp}
Assume that we role a die once:
\begin{enumerate}
    \item A is the event of getting an even number.
        $$ P(A) = \frac{1}{2} $$
    \item B is the event of getting a number $>$ 3.
        $$ P(B) = \frac{1}{2} $$
\end{enumerate}    
Are A \& B Independent?
$$ P(A\cap B) = \frac{1}{3} $$
$$ P(A)\times P(B) = \frac{1}{4} $$
A \& B are not independent
\end{exmp}    
\begin{exmp}
In a box, there are 8 red and 3 green
balls. We choose two balls randomly.
Find the probability that:
\begin{enumerate}
    \item Both balls are red.
        $$ P(2red)= P(r1)P(r2|r1) = \frac{8}{11} \times \frac{7}{10} $$
    \item Both balls are green.
        $$ P(2green) = P(g1)P(g2|g1) = \frac{3}{11} \times \frac{2}{10}$$
    \item First ball is red, and the second ball is
        $$ P(green,red) =  P(r1) \times P(g2|r1) 
        = \frac{8}{11} \times \frac{3}{10}
        $$
\end{enumerate}
\end{exmp}    
\begin{exmp}
     One card is drawn from a standard deck, replaced, and then the second card
     is drawn.  Find the probability that the first card is a king, and the
     second one is not.
 $$ P(E) = P(k) \times P(k') = \frac{4}{52} \times \frac{48}{52} $$
 Note that the two events are independent.
\end{exmp}
\begin{exmp}
     A box contains black chips and white chips. A person selects 2 chips. If
     the probability of selecting a black chip and a white chip is 15/56, and
     the probability of selecting a black chip on the first draw is 3/8.  
     
     Find the probability of selecting the white chip on the second draw, given that
     the first chip selected was a black chip.
     $$ P(W|B) = \frac{P(B \cap W)}{P(B)} = \frac{15/56}{3/8} $$
\end{exmp}

\section{Total Probability}
Let $F$ be any event that is already occurred. Then, The sample space can be
represented as $S = F \cup F'$ where $ F \cap F' = \emptyset $
$$ P(E) = P(E \cap S) = P(E \cap (F \cup F')) = $$
$$ P((E \cap F) \cup (E \cap F') ) = P(E\cap F)  + P(E \cap F') =$$
$$ P(F) \times P(E|F) + P(F') \times P(E|F')$$
It allows one to express the probability of any event in terms 
of conditional probabilities
$$ P(E) = P(F).P(E|F) + P(F').P(E|F'); \ \ F \cup F' = S; \ \ F \cap F' = \emptyset $$

\begin{exmp}
There are two identical bottles. One bottle
contains 2 green balls and 1 red ball, the
other bottle contains 2 red balls.
A bottle is selected at random, and a ball
is drawn.
What is the probability that the ball is red?\\


Let B1 and B2 stand for the events that the first bottle and the seconed bottle 
were selected. Hence $P(B_1) = P(B_2) = 0.5$ The chances to draw a red ball from b1 are 
    $\frac{1}{3}$. for b2 the cahnces are 1.
    \[
        P(r)= P(B_1)P(r|B_1) + P(B_2)P(r|B_2) =
    \]
    \[
        \frac{1}{2} \times \frac{1}{3} + \frac{1}{2} \times \frac{2}{2} 
    \]
\end{exmp}
\section{Bayes' Rule}
 Suppose that S is represented as the union of
finite number “n” of disjoint events $F_j$ .
 Then the total probability rule can be generalized as follows:
 \[
     \displaystyle P(E) = \sum_{k=1}^n P(F_k).P(E|F_k)
 \]
 \[
     \bigcup_{k=1}^n F_k = S; \ \ F_i \cap F_j = \emptyset : i \neq j.
 \]

\begin{exmp}
There are two identical bottles. One bottle
contains 2 green balls and 1 red ball, the
other bottle contains 2 red balls.
A bottle is selected at random, and a ball
is drawn.
What is the probability that the ball is red?\\


Let B1 and B2 stand for the events that the first bottle and the seconed bottle 
were selected. Hence $P(B_1) = P(B_2) = 0.5$ The chances to draw a red ball from b1 are 
    $\frac{1}{3}$. for b2 the cahnces are 1.
    \[
        P(r)= P(B_1)P(r|B_1) + P(B_2)P(r|B_2) =
    \]
    \[
        \frac{1}{2} \times \frac{1}{3} + \frac{1}{2} \times \frac{2}{2} 
    \]
\end{exmp}


\end{document}
